---
title: "Fit lasso VAR to stock indices dataset"
output: html_notebook
---

```{r}
rm(list=ls())

# load required packages
library("reticulate")
library("tidyr")
library("dplyr")
library("BigVAR")
library("data.table")

# load numpy package
np <- import("numpy")

# load sample dataset
npz_obj <- np$load(file.path(dirname(getwd()), "data", "world_stock_indexes", "americas_stock_indexes.npz"))

data <- npz_obj["X_np"]
k <- dim(data)[2]
p <- 5
penalty <- "BasicEN"
cv_type <- "Rolling"
steps <- 1
verbose <- FALSE
intercept <- FALSE
T = 100
```

To build the hyperparameter search procedure for the lambda value of the penalized VAR model, we use the "constructModel" function. This function has the following useful parameters:

> 1) gran: Two options for the grid of penalty parameters, the first option controls the depth of the lambda grid (a good default option is 50). The second option controls the number of grid values (a good default is 10);
> 2) h: Forecast horizon in which to optimize (default 1);
> 3) cv: Type of validation;
> 4) IC: Boolean. If True, returns AIC and BIC for the models;
> 5)  model.controls: ;


```{r}
model_search <- constructModel(Y = data,
                               p = p,
                               struct = penalty,
                               gran = c(150, 20), 
                               h = steps, 
                               cv = cv_type,
                               verbose = verbose,
                               IC = TRUE,
                               model.controls=list(intercept=FALSE))
results_model_search <- cv.BigVAR(model_search)
# results_model_search@LambdaGrid
results_model_search@OptimalLambda
```

Once we have the optimal lambda, we can fit the model in the data to produce the sparse matrix of coefficients:

```{r}
B <- BigVAR.fit(Y = data,
                struct = 'BasicEN',
                p = 5,
                lambda = results_model_search@OptimalLambda,
                intercept = FALSE)[,,1]
B <- B[ , 2:ncol(B)]
```

Finally, we can use the sparse matrix of coefficients to simulate similar VAR processes:

```{r}
A <- VarptoVar1MC(B, p, k)
Y <- MultVarSim(k = k,
                A1 = A,
                p = p,
                Sigma = .1*diag(k),
                T = T)

ts.plot(Y, col = 1:k)
```







